Cycles of the logistic map

### 2014-01-30 doc/pc-ijbc.tex
small corrections to the papers and web pages

\begin{abstract}
The onset and bifurcation points of the $n$-cycles of
  several polynomial maps are located
  through a characteristic equation
  connecting cyclic polynomials of the cycle points.
The polynomials satisfied by the parameters
  of the logistic, H\'enon, and cubic maps
  at the onset and bifurcation points are obtained
  for $n$ up to 14, 9, and 9,
  respectively.
\end{abstract}

@@

A fixed point is a solution of $x^* = f(x^*)$.
%
A stable fixed point requires that a small deviation from $x^*$
  gets reduced in magnitude after an iteration of \refeq{logmap}.
%
Since $f$ is differentiable,
  the condition means $|f'(x^*)| \le 1$
  \cite{strogatz, hao}.
%
An $n$-cycle is a sequence $x_1, \dots, x_n$
  such that $n$ is the smallest positive integer
  satisfying $x_1 = x_{n+1} = f^n(x_1)$,
  where $f^n$ is the $n$th iterate of $f$,
  e.g., $f^3(x) = f(f(f(x)))$.
By definition, any $x_k$ in an $n$-cycle of $f$ is a fixed point of $f^n$
(whereas the latter also permits an $x_k$ in any $d$-cycle with $d|n$).
%
%\big[but the converse is untrue, for a fixed point of $f^n$
%  can also be a fixed point of $f^d$ for a divisor $d$ of $n$:
%  if $f^d(x) = x$, then $f^n(x) = f^d(\cdots f^d(x)\cdots) = x$\big].
%
%
Thus, the stability of a cycle can be defined
  according to that of the corresponding fixed point of $f^n$:
  a stable cycle requires
  $\big| \frac {d} {dx_1} f^n(x_1) \big| \le 1$,
  or by the chain rule,
%
%
%
\begin{equation}
  \Big| f'(x_n) \cdots f'(x_1) \Big| \le 1,
\label{eq:der}
\end{equation}
%
%
The onset and bifurcation points
  are defined at the loci
  where $\frac {d} {dx_1} f^n(x_1)$ reaches $+1$ and $-1$,
  respectively \cite{strogatz, hao}.
%
%
%
The cycle and its stability depend on the parameter $r$,
%
and a stable cycle corresponds to a window of $r$
  between the onset point $r_a$ and bifurcation point $r_b$.
%
There are generally multiple such windows for a single $n$,
  but all $r_a$ are zeros of a polynomial,
  and all $r_b$ zeros of another.
%
Our aim is thus to find the two polynomials for a given $n$.
%
The problem was tackled either by hand for small $n$
\cite{brown1, brown2, stephenson1, stephenson2, stephenson3,
      saha, bechhoefer, gordon, burm, zhang}
or by general computer algebra methods \cite{bailey1, bailey2, kk1, lewis}.
%
Below we present a specialized algorithm, which offers some performance gain.
%
%
%
%The window boundaries can be classified
%  to onset and bifurcation points according to \refeq{der};
%Here, if $r > 0$, then
%  $r_a$ ($r_b$) are the onset (bifurcation) points.
%

@@

To simplify the calculation, we first change variables \cite{mandelbrot, brown1} as
%
%\begin{equation}
\[
    x^{\mathrm{new}} \equiv r(x^{\mathrm{old}} - 1/2),
    \quad R \equiv r(r-2)/4,
\]
%\label{eq:chvars}
%\end{equation}
%
and rewrite the map, in terms of $x^{\mathrm{new}}$, as
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv R - {x_k}^2.
  %\tag{$\ref{eq:logmap}'$}
\label{eq:logmaps}
\end{equation}
%
%\refeq{logmaps} simplifies \refeq{logmap} as $x_k$ now only occurs once.
%
We will solve the cycle boundaries as the zeros of some polynomials of $R$,
and the corresponding polynomials for $r$ can be obtained by $R \rightarrow r(r-2)/4$.
%
%the numerical $r$ values can always be obtained from $r = 1 + \sqrt{1 + 4R}$.
%

@@

Since the $n$-cycles points constitute
  a subset of the fixed points of $f^n$,
we will first find
  the polynomials of marginal stability
  for the fixed points of $f^n$
  (\refsecs{cyclic}-\ref{sec:examples}).
The polynomials contain
  extraneous factors contributed by short $d$-cycles with $d|n$,
  which are also fixed points of $f^n$.
The factors are removed in \refsec{primfac}.

@@

We call a cyclic polynomial simple
  if it is equal to the sum of
  all distinct cyclic versions of
  a unit-coefficient monomial of $x_k$.
%
The monomial is called a generator of the cyclic polynomial.
%
For example, the above $a$ and $b$ are simple,
  and $x_1 x_2$ and $x_1 x_3$ are their generators, respectively.
%
A simple cyclic polynomial
  $C_p(\vx{n})$, with $\vx{n} \equiv \{x_1, \ldots, x_n\}$,
can be labeled as follows.
%
If the monomial generator is
  ${x_1}^{e_1} {x_2}^{e_2} \dots {x_n}^{e_n}$,
  we define the label $p$ as the sequence of indices with
  $e_1$ 1's, $e_2$ 2's, \ldots, $e_n$ $n$'s.
%
For example,
$C_{12}(\vx{n})  = x_1 x_2 + x_2 x_3 + \dots + x_n x_1$
and
$C_{112}(\vx{n}) = {x_1}^2 x_2 + {x_2}^2 x_3 + \dots + {x_n}^2 x_1$.
%
Below we may drop %the subscript ``$_n$'' or even the argument
  ``$(\vx{n})$'' when $n$ is clear from the context.
In case the cyclic polynomial has several generators,
  we pick the one with the smallest label $p$
  in the sense of lexicographic order,
  e.g., we choose $x_1 x_2$ over $x_2 x_3$ as the generator of $C_{12}(\vx{n})$,
  because ``$12$'' $<$ ``$23$.''
%
Finally, we add $C_0(\vx{n}) \equiv 1$.
%
%
%

@@

A cyclic polynomial is square-free
  if it has no square or higher powers of $x_k$,
e.g., $C_{12} = x_1 x_2 + \cdots$ is square-free,
  $C_{112} = {x_1}^2 x_2 + \cdots$ is not.
The label $p$ of a simple square-free cyclic polynomial
  has no repeated index.
We denote the set of labels $p$ of all simple square-free cyclic polynomials by
$\Pset = \{0, 1, 12, 13, \ldots, 123, 124, \ldots, 12\cdots n\}$,
and its size by $\NB$ [cf. \refeq{necklace} below for the formula].
% is the number of square-free cyclic polynomials.
%
%
The set of the simple square-free cyclic polynomials
  furnishes a basis for expanding an arbitrary cyclic polynomial:

\begin{lemma}
  For the logistic map \refeq{logmaps},
  any cyclic polynomial $K(\vx{n})$
  formed by the $n$-cycle points
  $\vx{n} = \{x_1, \ldots, x_n\}$
  is a linear combination of
  the simple square-free cyclic polynomials $C_p(\vx{n})$:
\[
  K(\vx{n}) = \sum_{p \in \Pset} f_p(R) \, C_p(\vx{n}),
\]
  where $\Pset$ % = \{0, 1, 12, 13, \ldots, 12\dots n\}$ is
  is the set of labels of all simple square-free cyclic polynomials,
  and $f_p(R)$ are polynomials of $R$.
  \label{thm:sqrfree}
\end{lemma}
%
%
\begin{proof}
Consider the following square-free reduction.
%
Given a cyclic polynomial $K(\vx{n})$,
  we recursively apply \refeq{logmaps} as
  $x_k^2 \rightarrow R - x_{k+1}$,
  until all squares or higher powers of $x_k$ are eliminated.
The process will not last indefinitely
  for each substitution reduces the degree in $x_k$ %(no matter which $k$)
  by one.
Since both the replacement and the original polynomial are cyclic,
  so is the final square-free polynomial.
Terms involving no $x_k$
  are collected to serve as the coefficient before $C_0(\vx{n}) \equiv 1$.
The coefficients $f_p$ are polynomials of $R$,
  for $R$ is the only new variable introduced by the substitutions.
\end{proof}
%
%
For example, the cyclic polynomial
  $K(\vx{2}) = {x_1}^2 \, x_2 + {x_2}^2 \, x_1$
can be written as $K = (R + 1) \, C_{1} - 2 R \, C_0$
for
${x_1}^2 \, x_2 = R \, x_2 - {x_2}^2 = R \, x_2 - R + x_1$ and
${x_2}^2 \, x_1 = R \, x_1 - R + x_2$.

\reflem{sqrfree}
  shows that any cyclic polynomial can be expanded
  as a combination of the simple square-free ones.
The key step of the solution is to further show that
  at the onset and bifurcation points,
  the $\NB$ simple square-free cyclic polynomials $C_p$ are themselves
  linearly connected by a characteristic equation
  furnished by \refeq{der}.

@@

By \refeq{der},
  $\Lambda(\vx{n})$ equals $\lambda = +1$ ($-1$)
  at the onset (bifurcation) point,
  respectively;
  so \refeq{xcp0}
  becomes a homogeneous linear equation of
  $C_p(\vx{n})$:
\begin{equation}
  \lambda \, C_p(\vx{n}) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx{n}),
\label{eq:xcp}
\end{equation}
%
%
%where $\lambda$ is $+1$ and $-1$ at the onset and bifurcation points, respectively;
or in matrix form,
\begin{equation}
  \big[ \vct T(R) - \lambda \, \vct I \big] \, \vct C = 0,
\tag{\ref{eq:xcp}$'$}
\end{equation}
%
%
where
$\vct I$ is the $\NB \times \NB$ identity matrix,
$\vct T(R) = \{T_{pq}(R)\}$ is an $\NB \times \NB$ matrix,
and
$\vct C = \{C_p(\vx{n})\}$ is an $\NB$-dimensional column vector.

@@

The characteristic polynomial $A_n\Rlam$ in \refeq{xr}
  is derived for the fixed points of $f^n$.
Thus, it encompasses not just the $n$-cycles,
  but $d$-cycles for all divisors $d$ of $n$.
%
Fortunately, contributions from cycles of different periods
  are separate factors of $A_n\Rlam$.
For a large $n$,
  the factor $P_n\Rlam$ of $n$-cycles only,
  or the \emph{$n$-cycle polynomial} below,
  can be recognized
  as the largest irreducible factor of $A_n\Rlam$.
%
More systematically, we have
%\begin{theorem}
%  The polynomial $P_n\Rlam$ of all $n$-cycles
%  is a factor of the characteristic polynomial $A_n\Rlam$,
%  and can be computed as
  \begin{equation}
    P_n\Rlam
    = \prod_{cd = n} B_d^c\Rlam^{\mu(c)},
  \label{eq:primfac}
  \end{equation}
where
  $B_d^c\Rlam \equiv \prod_{k=1}^c A_{d}(R, e^{2k\pi i/c} \lambda^{1/c})$
  with
  $\lambda^{1/c}$ being a complex $c$th root of $\lambda$.
Particularly, $B_n^1\rlam = A_n\rlam$.
The outer product runs over all divisors $d$ of $n$.
%\end{theorem}
%
%
%\begin{remark}
%Here, ``minimal'' means that the polynomial contains all zeros.
%It happens to be irreducible for a general $\lambda$.
%At $\lambda = 1$, the polynomial of $R$, however, can be factorized (see later).
%\end{remark}
%
%\begin{remark}
The M\"obius function $\mu(n)$ is $(-1)^k$
  if $n$ is the product of $k$ distinct primes,
  or 0 if $n$ is divisible by a square of a prime:
%Thus,
$\mu(n) = 1, -1, -1, 0, -1, 1, -1, \ldots,$
  from $n = 1$
%The $\mu(n)$ is useful for inversion:
%$g(n) = \sum_{d|n} \mu(n/d) \, h(d)$
%if and only if $h(n) = \sum_{d|n} g(d)$
 \cite{hardy}.
%\end{remark}
%

%\begin{remark}
%$B_{d,c}\lambda$ is a polynomial of $\lambda$.
Despite the radical argument $\lambda^{1/c}$,
  the product
  $B_d^c\Rlam = \prod_{k=1}^c A_d(R, e^{2k\pi i/c} \lambda^{1/c})$
  is a polynomial of $\lambda$,
  for it is invariant under
  $\lambda \rightarrow e^{2\pi i} \lambda$.
Also, $\deg_\lambda B_d^c\Rlam = \deg_\lambda A_d\Rlam$.
For convenience, we will drop the $R$ in ``$\Rlam$'' %or the arguments entirely below
  till the end the next section.
%
%\end{remark}

@@

We coded the algorithm into a Mathematica program,
  and computed the polynomials for $n$ up to 14.
The data are saved on the website http://logperiod.appspot.com.
%
%The bifurcation polynomial of the 14-cycle in shown in \reffig{T14b}.
%
%The exact polynomials of these maps are generally too large to print on paper,
%  e.g., the bifurcation polynomial of the 14-cycle
%    takes about 27 megabytes.
%
%
%Compared with other approaches \cite{hitzl, gordon, bailey2, kk1, hao, hao2}
%it appeared to be simpler and more transparent,
%as it only requires the computation of a determinant.
%
As shown in \reffig{logtime}, the algorithm
  outperformed the resultant \cite{burm} and
  Gr\"obner basis \cite{kk1} methods.
%  by at least two orders of magnitudes.
For the implementation of the resultant method on Magma, we have included a version
  that used the polynomial interpolation technique
  similar to the one described in the previous paragraph.
This version performed slightly better than the purely symbolic version.
Since the resultant contains
  only the $n$th power of the $n$-cycle polynomial $P_n\Rlam$,
the number of evaluation points needs to be doubled
  to recover the sign of $P_n\Rlam$ for an even $n$.
%
The current method works better than the resultant method
  because the resultant of $f^n(x)-x$ and $\frac{d}{dx}f^n(x)-\lambda$
  is the determinant of the Sylvester matrix, which has
  $2^{n+1}-1$ rows and columns
  [as $\deg_x f^n(x) = 2^n$],
  while the coefficient matrix in \refeq{xr}
  has only $\NB \approx 2^n/n$ [cf. \refeq{necklace}] rows and columns.
%Thus, the current algorithm is expected to be faster.

@@

The enlarged basis significantly increases the dimension of
  the coefficient matrix in \refeq{xr}, hence the computation cost.
%
Below we discuss two techniques to alleviate the problem.
%by exploiting the certain properties of the map.
%First the antisymmetry: $f_r(-x) = -f_r(x)$
%
First, we may include in the basis
  about only half of the simple cyclic polynomials,
  that is, those of \emph{even} degrees in $\vx{n}$.
%
This is because \refeq{cubreplace} %contains only linear and cubic terms,
  never reduces even-degree polynomials to odd-degree ones,
  or the other way around.
%
The characteristic equation obtained
  from the basis of even-degree cyclic polynomials
  is the same as that
  from the basis of odd-degree cyclic polynomials for an odd $n$.
But for an even $n$, the former admits several symmetric cycles
  that are absent in the latter.
The difference is due to the symmetry of \refeq{cubic},
  and is discussed below.

@@

To be more precise, we first define the $n$-half-cycle polynomial $H_n\rlam$
  in analogous to $P_n\rlam$.
%
It is convenient to exclude the contribution
  from the trivial fixed point at $x = 0$
  by defining the \emph{nontrivial} $n$-cycle polynomial
  $P^*_n\rlam \equiv P_n\rlam/(r-\lambda)^{\delta_{n,1}}$,
  and %the nontrivial
  $n$-half-cycle polynomial
  $H^*_n\rlam \equiv H_n\rlam/(r-\lambda)^{\delta_{n,1}}$.
%
%Since a repeated $n$-half-cycle is a $2n$-cycle,
Dividing the $2n$-cycle polynomial
  $P^{*}_{2n}(r, \lambda)$
  by
  $H^{*}_{n}\big(r, \sqrt \lambda\big)
   H^{*}_{n}\big(r, -\sqrt \lambda\big)$
%
yields the \emph{asymmetric} $n$-cycle polynomial:
\begin{equation}
\tilde P^*_n\rlam \equiv
  %\left.
  \frac{
    P^*_n\rlam
  }
  %\middle/
  {
    \left[
    \prod_{\sigma = \pm 1}
      H^*_{n/2}(r, \sigma \sqrt\lambda)
%      H^*_{n/2}(r, -\sqrt\lambda)
    \right]^{\even(n)}
  },
  %\right.
\label{eq:Pasym}
\end{equation}
%]
where $\even(n) \equiv 1 - n\bmod 2$.
%
One can verify that $\tilde P^*_n\rlam$ is the same as the $n$-cycle polynomial
  derived from the basis of odd-degree cyclic polynomials.
%

>>>>

\begin{abstract}
The onset and bifurcation points of $n$-cycles of
  several polynomial maps are located
  through a characteristic equation
  connecting cyclic polynomials of the cycle points.
The polynomials satisfied by the parameters
  of the logistic, H\'enon, and cubic maps
  at the onset and bifurcation points are obtained
  for $n$ up to 14, 9, and 9,
  respectively.
\end{abstract}

@@

A fixed point is a solution of $x^* = f(x^*)$.
%
A stable fixed point requires that a small deviation from $x^*$
  gets reduced in magnitude after an iteration of \refeq{logmap}.
%
Since $f$ is differentiable,
  the condition means $|f'(x^*)| \le 1$
  \cite{strogatz, hao}.
%
An $n$-cycle is a sequence $x_1, \dots, x_n$
  such that $n$ is the smallest positive integer
  satisfying $x_1 = x_{n+1} = f^n(x_1)$,
  where $f^n$ is the $n$th iterate of $f$,
  e.g., $f^3(x) = f(f(f(x)))$.
By definition, any $x_k$ in an $n$-cycle of $f$ is a fixed point of $f^n$
(whereas the latter also permits an $x_k$ in any $d$-cycle with $d|n$).
%
%\big[but the converse is untrue, for a fixed point of $f^n$
%  can also be a fixed point of $f^d$ for a divisor $d$ of $n$:
%  if $f^d(x) = x$, then $f^n(x) = f^d(\cdots f^d(x)\cdots) = x$\big].
%
%
Thus, the stability of a cycle can be defined
  according to that of the corresponding fixed point of $f^n$:
  a stable cycle requires
  $\big| \frac {d} {dx_1} f^n(x_1) \big| \le 1$,
  or by the chain rule,
%
%
%
\begin{equation}
  \Big| f'(x_n) \cdots f'(x_1) \Big| \le 1,
\label{eq:der}
\end{equation}
%
%
The onset and bifurcation points
  are defined at the loci
  where $\frac {d} {dx_1} f^n(x_1)$ reaches $+1$ and $-1$,
  respectively \cite{strogatz, hao}.
%
%
%
The cycle and its stability depend on the parameter $r$,
%
and a stable cycle corresponds to a window of $r$
  between the onset point $r_a$ and bifurcation point $r_b$.
%
There are generally multiple such windows for a single $n$,
  but all $r_a$ are zeros of a polynomial,
  and all $r_b$ zeros of another.
%
Our aim is thus to find the two polynomials for a given $n$.
%
The problem was tackled either by hand for small $n$
\cite{brown1, brown2, stephenson1, stephenson2, stephenson3,
      saha, bechhoefer, gordon, burm, zhang}
or by general computer-algebraic methods \cite{bailey1, bailey2, kk1, lewis}.
%
Below we present a specialized algorithm, which offers some performance gain.
%
%
%
%The window boundaries can be classified
%  to onset and bifurcation points according to \refeq{der};
%Here, if $r > 0$, then
%  $r_a$ ($r_b$) are the onset (bifurcation) points.
%

@@

To simplify the calculation, we first change variables \cite{mandelbrot, brown1} as
%
%\begin{equation}
\[
    x^{\mathrm{new}} \equiv r(x^{\mathrm{old}} - 1/2),
    \quad R \equiv r(r-2)/4,
\]
%\label{eq:chvars}
%\end{equation}
%
and rewrite the map, in terms of $x^{\mathrm{new}}$, as
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv R - {x_k}^2.
  %\tag{$\ref{eq:logmap}'$}
\label{eq:logmaps}
\end{equation}
%
%\refeq{logmaps} simplifies \refeq{logmap} as $x_k$ now only occurs once.
%
We will solve the cycle boundaries as zeros of some polynomials of $R$,
and the corresponding polynomials for $r$ can be obtained by $R \rightarrow r(r-2)/4$.
%
%the numerical $r$ values can always be obtained from $r = 1 + \sqrt{1 + 4R}$.
%

@@

Since the $n$-cycles points constitute
  a subset of the fixed points of $f^n$,
we will first find
  the polynomials of marginal stability
  for the fixed points of $f^n$
  (\refsecs{cyclic}-\ref{sec:examples}).
The polynomials contain
  extraneous factors contributed by shorter $d$-cycles with $d|n$,
  which are also fixed points of $f^n$.
The factors are removed in \refsec{primfac}.

@@

We call a cyclic polynomial simple
  if it is equal to the sum of
  all distinct cyclic versions of
  a unit-coefficient monomial of $x_k$.
%
The monomial is called a generator of the cyclic polynomial.
%
For example, the above $a$ and $b$ are simple,
  and $x_1 x_2$ and $x_1 x_3$ are their generators, respectively.
%
A simple cyclic polynomial
  $C_p(\vx{n})$, with $\vx{n} \equiv \{x_1, \ldots, x_n\}$,
can be labeled as follows.
%
If the monomial generator is
  ${x_1}^{e_1} {x_2}^{e_2} \dots {x_n}^{e_n}$,
  we define the label $p$ as the sequence of indices with
  $e_1$ 1's, $e_2$ 2's, \ldots, $e_n$ $n$'s.
%
For example,
$C_{12}(\vx{n})  = x_1 x_2 + x_2 x_3 + \dots + x_n x_1$
and
$C_{112}(\vx{n}) = {x_1}^2 x_2 + {x_2}^2 x_3 + \dots + {x_n}^2 x_1$.
%
Below we may drop %the subscript ``$_n$'' or even the argument
  ``$(\vx{n})$'' when $n$ is clear from the context.
In case the cyclic polynomial has several generators,
  we pick the one with the smallest label $p$
  in lexicographic order,
  e.g., we choose $x_1 x_2$ over $x_2 x_3$ as the generator of $C_{12}(\vx{n})$,
  because ``$12$'' $<$ ``$23$.''
%
Finally, we add $C_0(\vx{n}) \equiv 1$.
%
%
%

@@

A cyclic polynomial is square-free
  if it has no square or higher powers of $x_k$,
e.g., $C_{12} = x_1 x_2 + \cdots$ is square-free,
  $C_{112} = {x_1}^2 x_2 + \cdots$ is not.
The label $p$ of a simple square-free cyclic polynomial
  has no repeated index.
We denote the set of labels $p$ of all simple square-free cyclic polynomials by
$\Pset = \{0, 1, 12, 13, \ldots, 123, 124, \ldots, 12\cdots n\}$,
and its size by $\NB$ [cf. \refeq{necklace} below for the formula].
% is the number of square-free cyclic polynomials.
%
%
The set of simple square-free cyclic polynomials
  furnishes a basis for expanding an arbitrary cyclic polynomial:

\begin{lemma}
  For the logistic map \refeq{logmaps},
  any cyclic polynomial $K(\vx{n})$
  formed by the $n$-cycle points
  $\vx{n} = \{x_1, \ldots, x_n\}$
  is a linear combination of
  the simple square-free cyclic polynomials $C_p(\vx{n})$:
\[
  K(\vx{n}) = \sum_{p \in \Pset} f_p(R) \, C_p(\vx{n}),
\]
  where $\Pset$ % = \{0, 1, 12, 13, \ldots, 12\dots n\}$ is
  is the set of labels of all simple square-free cyclic polynomials,
  and $f_p(R)$ are polynomials of $R$.
  \label{thm:sqrfree}
\end{lemma}
%
%
\begin{proof}
Consider the following square-free reduction.
%
Given a cyclic polynomial $K(\vx{n})$,
  we recursively apply \refeq{logmaps} as
  ${x_k}^2 \rightarrow R - x_{k+1}$,
  until all squares or higher powers of $x_k$ are eliminated.
The process will not last indefinitely
  for each substitution reduces the degree in $x_k$ %(no matter which $k$)
  by one.
Since both the replacement and the original polynomial are cyclic,
  so is the final square-free polynomial.
Terms involving no $x_k$
  are collected to serve as the coefficient before $C_0(\vx{n}) \equiv 1$.
The coefficients $f_p$ are polynomials of $R$,
  for $R$ is the only new variable introduced by the substitutions.
\end{proof}
%
%
For example, the cyclic polynomial
  $K(\vx{2}) = {x_1}^2 \, x_2 + {x_2}^2 \, x_1$
can be written as $K = (R + 1) \, C_{1} - 2 R \, C_0$
for
${x_1}^2 \, x_2 = R \, x_2 - {x_2}^2 = R \, x_2 - R + x_1$ and
${x_2}^2 \, x_1 = R \, x_1 - R + x_2$.

\reflem{sqrfree}
  shows that any cyclic polynomial can be expanded
  as a combination of simple square-free ones.
The key step of the solution is to further show that
  at the onset and bifurcation points,
  the $\NB$ simple square-free cyclic polynomials $C_p$ are themselves
  linearly connected by a characteristic equation
  furnished by \refeq{der}.

@@

By \refeq{der},
  $\Lambda(\vx{n})$ equals $\lambda = +1$ ($-1$)
  at the onset (bifurcation) point;
  so \refeq{xcp0}
  becomes a homogeneous linear equation of
  $C_p(\vx{n})$:
\begin{equation}
  \lambda \, C_p(\vx{n}) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx{n}),
\label{eq:xcp}
\end{equation}
%
%
%where $\lambda$ is $+1$ and $-1$ at the onset and bifurcation points, respectively;
or in matrix form,
\begin{equation}
  \big[ \vct T(R) - \lambda \, \vct I \big] \, \vct C = 0,
\tag{\ref{eq:xcp}$'$}
\end{equation}
%
%
where
$\vct I$ is the $\NB \times \NB$ identity matrix,
$\vct T(R) = \{T_{pq}(R)\}$ is an $\NB \times \NB$ matrix,
and
$\vct C = \{C_p(\vx{n})\}$ is an $\NB$-dimensional column vector.

@@

The characteristic polynomial $A_n\Rlam$ in \refeq{xr}
  is derived for the fixed points of $f^n$.
Thus, it encompasses not just the $n$-cycles,
  but $d$-cycles for all divisors $d$ of $n$.
%
Fortunately, contributions from cycles of different periods
  are separate factors of $A_n\Rlam$.
For a large $n$,
  the factor $P_n\Rlam$ of the $n$-cycles only,
  or the \emph{$n$-cycle polynomial} below,
  can be recognized
  as the largest irreducible factor of $A_n\Rlam$.
%
More systematically, we have
%\begin{theorem}
%  The polynomial $P_n\Rlam$ of all $n$-cycles
%  is a factor of the characteristic polynomial $A_n\Rlam$,
%  and can be computed as
  \begin{equation}
    P_n\Rlam
    = \prod_{cd = n} B_d^c\Rlam^{\mu(c)},
  \label{eq:primfac}
  \end{equation}
where
  $B_d^c\Rlam \equiv \prod_{k=1}^c A_{d}(R, e^{2k\pi i/c} \lambda^{1/c})$
  with
  $\lambda^{1/c}$ being a complex $c$th root of $\lambda$.
Particularly, $B_n^1\rlam = A_n\rlam$.
The outer product runs over all divisors $d$ of $n$.
%\end{theorem}
%
%
%\begin{remark}
%Here, ``minimal'' means that the polynomial contains all zeros.
%It happens to be irreducible for a general $\lambda$.
%At $\lambda = 1$, the polynomial of $R$, however, can be factorized (see later).
%\end{remark}
%
%\begin{remark}
The M\"obius function $\mu(n)$ is $(-1)^k$
  if $n$ is the product of $k$ distinct primes,
  or 0 if $n$ is divisible by a square of a prime:
%Thus,
$\mu(n) = 1, -1, -1, 0, -1, 1, -1, \ldots,$
  from $n = 1$
%The $\mu(n)$ is useful for inversion:
%$g(n) = \sum_{d|n} \mu(n/d) \, h(d)$
%if and only if $h(n) = \sum_{d|n} g(d)$
 \cite{hardy}.
%\end{remark}
%

%\begin{remark}
%$B_{d,c}\lambda$ is a polynomial of $\lambda$.
Despite the radical argument $\lambda^{1/c}$,
  the product
  $B_d^c\Rlam = \prod_{k=1}^c A_d(R, e^{2k\pi i/c} \lambda^{1/c})$
  is a polynomial of $\lambda$,
  for it is invariant under
  $\lambda \rightarrow e^{2\pi i} \lambda$.
Also, $\deg_\lambda B_d^c\Rlam = \deg_\lambda A_d\Rlam$.
For convenience, we will drop the $R$ in ``$\Rlam$'' %or the arguments entirely below
  till the end of the next section.
%
%\end{remark}

@@

We coded the algorithm into a Mathematica program,
  and computed the polynomials for $n$ up to 14.
The data are saved on the website http://logperiod.appspot.com.
%
%The bifurcation polynomial of the 14-cycle in shown in \reffig{T14b}.
%
%The exact polynomials of these maps are generally too large to print on paper,
%  e.g., the bifurcation polynomial of the 14-cycle
%    takes about 27 megabytes.
%
%
%Compared with other approaches \cite{hitzl, gordon, bailey2, kk1, hao, hao2}
%it appeared to be simpler and more transparent,
%as it only requires the computation of a determinant.
%
As shown in \reffig{logtime}, the algorithm
  outperformed the resultant \cite{burm} and
  Gr\"obner basis \cite{kk1} methods.
%  by at least two orders of magnitudes.
In implementing the resultant method on Magma, we have coded a version
  that used the polynomial-interpolation technique
  similar to the one described in the previous paragraph.
This version performed slightly better than the purely symbolic version.
Since the resultant contains
  only the $n$th power of the $n$-cycle polynomial $P_n\Rlam$,
the number of evaluation points needs to be doubled
  to recover the sign of $P_n\Rlam$ for an even $n$.
%
The current method works better than the resultant method
  because the resultant of $f^n(x)-x$ and $\frac{d}{dx}f^n(x)-\lambda$
  is the determinant of the Sylvester matrix, which has
  $2^{n+1}-1$ rows and columns
  [as $\deg_x f^n(x) = 2^n$],
  while the coefficient matrix in \refeq{xr}
  has only $\NB \approx 2^n/n$ [cf. \refeq{necklace}] rows and columns.
%Thus, the current algorithm is expected to be faster.

@@

The enlarged basis significantly increases the dimension of
  the coefficient matrix in \refeq{xr}, hence the computational cost.
%
Below we discuss two techniques to alleviate the problem.
%by exploiting the certain properties of the map.
%First the antisymmetry: $f_r(-x) = -f_r(x)$
%
First, we may include in the basis
  about only half of the simple cyclic polynomials,
  that is, those of \emph{even} degrees in $\vx{n}$.
%
This is because \refeq{cubreplace} %contains only linear and cubic terms,
  never reduces even-degree polynomials to odd-degree ones,
  or the other way around.
%
The characteristic equation obtained
  from the basis of even-degree cyclic polynomials
  is the same as that
  from the basis of odd-degree cyclic polynomials for an odd $n$.
But for an even $n$, the former admits several symmetric cycles
  that are absent in the latter.
The difference is due to the symmetry of \refeq{cubic},
  and is discussed below.

@@

To be more precise, we first define the $n$-half-cycle polynomial $H_n\rlam$
  analogous to $P_n\rlam$.
%
It is convenient to exclude the contribution
  from the trivial fixed point at $x = 0$
  by defining the \emph{nontrivial} $n$-cycle polynomial
  $P^*_n\rlam \equiv P_n\rlam/(r-\lambda)^{\delta_{n,1}}$,
  and %the nontrivial
  $n$-half-cycle polynomial
  $H^*_n\rlam \equiv H_n\rlam/(r-\lambda)^{\delta_{n,1}}$.
%
%Since a repeated $n$-half-cycle is a $2n$-cycle,
Dividing the $2n$-cycle polynomial
  $P^{*}_{2n}(r, \lambda)$
  by
  $H^{*}_{n}\big(r, \sqrt \lambda\big)
   H^{*}_{n}\big(r, -\sqrt \lambda\big)$
%
yields the \emph{asymmetric} $n$-cycle polynomial:
\begin{equation}
\tilde P^*_n\rlam \equiv
  %\left.
  \frac{
    P^*_n\rlam
  }
  %\middle/
  {
    \left[
    \prod_{\sigma = \pm 1}
      H^*_{n/2}(r, \sigma \sqrt\lambda)
%      H^*_{n/2}(r, -\sqrt\lambda)
    \right]^{\even(n)}
  },
  %\right.
\label{eq:Pasym}
\end{equation}
%]
where $\even(n) \equiv 1 - n\bmod 2$.
%
One can verify that $\tilde P^*_n\rlam$ is the same as the $n$-cycle polynomial
  derived from the basis of odd-degree cyclic polynomials.
%

We now calculate the degrees of the above polynomials
  to facilitate the polynomial-interpolation technique
  mentioned in \refsec{end}.
The number of even-degree cube-free cyclic polynomials
is equal to that of \emph{even ternary} necklaces,
in which each bead represents a number 0, 1, or 2,
  with the total being even.
This number is given by
%
\begin{equation}
  N_e(n) = 1 +
    \frac{1}{2n} \sum_{c d = n} \varphi(c)
    \left[ 2 - \odd(c) \, \right]
    (3^d-1),
\notag
%\label{eq:cubnecklace}
\end{equation}
where
$\odd(c) \equiv c \bmod 2$.
%$N_e(n) =$ 2, 4, 6, 14, 26, 68, 158, 424, \dots, from $n = 1$.
%
%This is roughly half of the number of all (even and odd) ternary necklaces:
%$N_3(n) = \frac{1}{n} \sum_{d|n} \varphi(n/d) 3^d \approx 2 N_e$.
%
%
%Since $N_e(n)$ is the degree in $\lambda$ of $A_n\rlam$,
%  half of the cyclic polynomials and their associated equations are spared.
%
The degree $L_e(n)$ of the $n$-cycle polynomial $P_n\rlam$
  can be obtained from
  $N_e(n) = \sum_{d|n} L_e(d)$,
and
%
%
\begin{equation}
  L_e(n) = \delta_{n, 1} +
    \frac{1}{2n} \sum_{cd = n} \mu(c) \odd(c) \, (3^d-1),
\notag
%\label{eq:cublyndon}
\end{equation}
%
%
%The first few values are 2, 2, 4, 10, 24, 60, 156, 410, \ldots, starting from $n = 1$.
%
which is also the number of $n$-cycles \cite{zeng},
%
and
  $L^*_e(n) \equiv \deg_\lambda P^*_n\rlam = L_e(n) - \delta_{n,1}$.
%
%For $n>1$, half of the $L^*_e(n)$ cycles
%  with negative $r$ are imaginary.
%They can, however, be made real by
%$x_k = \sqrt{r} z_k$;
%the transformed map
%$z_{k+1} = r \, z_k \, (1 - {z_k}^2)$
%shares the same boundary polynomials of $r$ with \refeq{cubic}.
%
%
One can further show that the degrees in $r$ are
  $\deg_r A_n\rlam = n N_e(n)$
  and
  $\deg_r P^{(*)}_n\rlam = n L^{(*)}_e(n)$.
